From Archives to AI: Residential Property Data Across Three Decades in Brunei Darussalam

### 2025-03-05 manuscript.qmd
Update manuscript

---
title: 'From Archives to AI: Residential Property Data Across Three Decades in Brunei Darussalam'
authors:
  - name: Haziq Jamil
    orcid: 0000-0003-3298-1010
    email: haziq.jamil@ubd.edu.bn
    url: 'https://haziqj.ml'
    corresponding: true
    affiliations:
      - name: Universiti Brunei Darussalam
        id: ubd
        department: Mathematical Sciences, Faculty of Science
        address: Universiti Brunei Darussalam, Jalan Tungku Link
        city: Bandar Seri Begawan
        country: Brunei
        postal-code: BE 1410
  - name: Amira Barizah Noorosmawie
    affiliations:
      - ref: ubd        
  - name: Hafeezul Waezz Rabu
    affiliations:
      - ref: ubd
  - name: Lutfi Abdul Razak
    orcid: 0000-0003-1681-0895
    email: lutfi.razak@ubd.edu.bn
    affiliations:
      - name: UBD School of Business and Economics
        id: ubdsbe
        # department: Mathematical Sciences, Faculty of Science
        address: Universiti Brunei Darussalam, Jalan Tungku Link
        city: Bandar Seri Begawan
        country: Brunei
        postal-code: BE 1410         
date-modified: last-modified
abstract: |
  This article introduces the first publicly available data set for analyzing the Brunei housing market, covering 31,150 property listings from 1993 to early 2025. The data set, curated from property advertisements in newspapers and online platforms, includes key attributes such as price, location, property type, and physical characteristics, enriched with area-level spatial information. Comprehensive and historical, it complements the Brunei Darussalam Central Bank’s Residential Property Price Index (RPPI), addressing the limitations of restricted access to raw RPPI data and its relatively short timeline since its inception in 2015. Data collection involved manual transcription from archival sources and automated web scraping using programmatic methods, supported by innovative cleaning with Large Language Models (LLMs) to structure unformatted text. The data set enables spatial and temporal analysis, with potential applications in economics, urban planning, and real estate research. While minor limitations exist, such as missing values and spatial coverage bias toward the Brunei-Muara district, this data set provides a robust foundation for exploring housing market trends and informing policy decisions.
keywords:
  - Housing Market
  - Brunei
  - Property Listings
  - Spatial Data
  - Web Scraping
  - Large Language Models
---

@@

- **Methodological innovation in data curation**. The data collection process employs a unique combination of manual transcription from archival sources, automated web scraping, and AI-based data cleaning using Large Language Models (LLMs). This open source and multi-method approach enhances reproducibility and transparency in data curation. The present work provides a reproducible framework for assembling complex datasets, which can be adapted for similar data collection efforts in other other domains, thereby advancing best practices in the field.

@@

- **Opportunities for handling of missing data**. The dataset includes instances of missing or inconsistent data, particularly in house characteristics, which present opportunities for further research on data imputation techniques. Researchers may explore advanced statistical and machine learning methods to address missing values and improve data reliability. The structured format and detailed documentation of the dataset support such methodological investigations, making it a valuable testbed for developing and evaluating new approaches to handling incomplete data in real-world datasets.

@@

:::: {.content-visible when-format="html"}
::: column-page
{{< embed notebooks/analysis.qmd#tbl-summary >}}
:::
::::

:::: {.content-visible unless-format="html"}
::: landscape
{{< embed notebooks/analysis.qmd#tbl-summary >}}
:::
::::

@@

These data validation procedures collectively contribute to a robust and reliable dataset, providing a solid foundation for subsequent analysis.

>>>>

---
title: 'From Archives to AI: Residential Property Data Across Three Decades in Brunei Darussalam'
authors:
  - name: Haziq Jamil
    orcid: 0000-0003-3298-1010
    email: haziq.jamil@ubd.edu.bn
    url: 'https://haziqj.ml'
    corresponding: true
    affiliations:
      - name: Universiti Brunei Darussalam
        id: ubd
        department: Mathematical Sciences, Faculty of Science
        address: Universiti Brunei Darussalam, Jalan Tungku Link
        city: Bandar Seri Begawan
        country: Brunei
        postal-code: BE 1410
  - name: Amira Barizah Noorosmawie
    affiliations:
      - ref: ubd        
  - name: Hafeezul Waezz Rabu
    affiliations:
      - ref: ubd
  - name: Lutfi Abdul Razak
    orcid: 0000-0003-1681-0895
    email: lutfi.razak@ubd.edu.bn
    affiliations:
      - name: UBD School of Business and Economics
        id: ubdsbe
        # department: Mathematical Sciences, Faculty of Science
        address: Universiti Brunei Darussalam, Jalan Tungku Link
        city: Bandar Seri Begawan
        country: Brunei
        postal-code: BE 1410         
date-modified: last-modified
abstract: |
  This article introduces the first publicly available data set for analyzing the Brunei housing market, covering more than 30,000 property listings from 1993 to early 2025. The data set, curated from property advertisements in newspapers and online platforms, includes key attributes such as price, location, property type, and physical characteristics, enriched with area-level spatial information. Comprehensive and historical, it complements the Brunei Darussalam Central Bank’s Residential Property Price Index (RPPI), addressing the limitations of restricted access to raw RPPI data and its relatively short timeline since its inception in 2015. Data collection involved manual transcription from archival sources and automated web scraping using programmatic methods, supported by innovative cleaning with Large Language Models (LLMs) to structure unformatted text. The data set enables spatial and temporal analysis, with potential applications in economics, urban planning, and real estate research. Although listing prices serve only as a proxy for market values and may deviate from actual sale prices due to negotiation dynamics and other factors, this data set still provides a valuable resource for reproducible analysis of housing market trends and for informing policy decisions.
keywords:
  - Housing Market
  - Brunei
  - Property Listings
  - Spatial Data
  - Web Scraping
  - Large Language Models
---

@@

- **Methodological innovation in data curation**. The data collection process employs a unique combination of manual transcription from archival sources, automated web scraping, and AI-based data cleaning using Large Language Models (LLMs). This open source and multi-method approach enhances reproducibility and transparency in data curation. The present work provides a reproducible framework for assembling complex data sets, which can be adapted for similar data collection efforts in other other domains, thereby advancing best practices in the field.

@@

- **Opportunities for handling of missing data**. The data set includes instances of missing or inconsistent data, particularly in house characteristics, which present opportunities for further research on data imputation techniques. Researchers may explore advanced statistical and machine learning methods to address missing values and improve data reliability. The structured format and detailed documentation of the data set support such methodological investigations, making it a valuable testbed for developing and evaluating new approaches to handling incomplete data in real-world data sets.

@@

```{r}
#| include: false
tbl_summary <- 
  hsp |>
  mutate(
    price = price / 1000,
    storeys = factor(
      ifelse(storeys >= 3, "3+", as.character(storeys)),
      levels = c("1", "2", "3+")
    ),
    district = gsub("Brunei Muara", "Brunei-Muara", district),
    district = factor(district, levels = c("Brunei-Muara", "Belait", "Tutong", "Temburong"))
  ) |>
  tbl_summary(
    include = c(price:baths),
    by = district,
    missing = "no", 
    type = all_continuous() ~ "continuous2",
    statistic = list(
      all_continuous2() ~ c("{mean} ({sd})", Range = "{min} - {max}", "{median} ({p25}, {p75})")
    ),
    label = list(
      price = "Price (BND 1,000)",
      type = "Property type",
      plot_area = "Plot area (acres)",
      floor_area = "Floor area (sq. ft.)",
      storeys = "Number of storeys",
      status = "Development status",
      tenure = "Land tenure",    
      beds = "Number of bedrooms",
      baths = "Number of bathrooms"
    ),
    digits = list(
      baths = 1,
      beds = 1
    )
  ) |>
  add_overall() |>
  add_n() |>
  modify_header(label ~ "**Variable**") |>
  as_gt()
```

:::: {.content-visible when-format="html"}
::: column-page
::: {#tbl-summary}
```{r}
#| html-table-processing: none
tbl_summary
```
Summary of housing data.
:::
:::
::::

:::: {.content-visible unless-format="html"}
::: landscape
::: {#tbl-summary}
```{r}
#| html-table-processing: none
tbl_summary |>
  tab_options(table.font.size = "9px")
```
Summary of housing data.
:::
:::
::::

@@

These data validation procedures collectively contribute to a robust and reliable data set, providing a solid foundation for subsequent analysis.
